### 2015-05-11 evaluation.tex
rewriten limitations

We performed experiments to validate the usefulness of the \whoop approach (see Section~\ref{whoop}). We first present race-checking results from running \whoop and \corral on \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel distribution\footnote{https://www.kernel.org}. We then evaluate the runtime performance and scalability of \corral and \whoop + \corral with different yield instrumentation granularities and context-switch bounds, to show that \whoop can efficiently accelerate bug-finding with \corral.

>>>>

We performed experiments to validate the usefulness of the \whoop approach (see Section~\ref{whoop}). We first present race-checking results from running \whoop and \corral on \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel distribution\footnote{\url{https://www.kernel.org}}. We then evaluate the runtime performance and scalability of \corral and \whoop + \corral with different yield instrumentation granularities and context-switch bounds, to show that \whoop can efficiently accelerate bug-finding with \corral.

### 2015-05-11 whoop.tex
rewriten limitations

The aim of this phase (see Figure~\ref{fig:whoop} -- A) is to take an input Linux driver written in C, together with an environmental model, and output an over-approximated version of the program in Boogie IVL, a simple imperative language with well-defined verification-focused semantics used as the input to a number of cutting-edge verifiers (e.g. Boogie and \corral). To achieve this translation, \whoop uses three LLVM\footnote{http://llvm.org}-based tools: Chauffeur\footnote{https://github.com/mc-imperial/chauffeur}, Clang\footnote{http://clang.llvm.org} and SMACK\footnote{https://github.com/smackers/smack}~\cite{rakamaric2014smack}.

@@

\whoop inherits all limitations of lockset analysis (as discussed in Section~\ref{bg:lockset}): a violation of the locking discipline does not always correspond to a real error (e.g. \whoop does not reason about lockfree data structures in a special way, it just over-approximates their read accesses); and we do not check for dynamically created locks (or for locks that are provided by external libraries). Because we perform two thread reduction, we also over-approximate the read accesses, which is another source of false positives.

Environmental modeling is another limitation of \whoop. As aforementioned, enriching \whoop with domain-specific knowledge to increase precision is an ongoing manual effort, but requires expertise in the Linux kernel. We argue that further increasing the precision of our models is orthogonal to the contributions of this paper. Moreover, despite unavoidable imprecisions during the sound static data race analysis, \whoop can use the reported information to significantly speedup precise bug-finders for concurrent programs, as seen in Sections~\ref{whoop:bugfinding} and~\ref{evaluation}.

Finally, \whoop is unable to verify drivers where the programmer implicitly enforces serialization of entry points that would otherwise race. For example, in the intel\_scu\_wd driver (see Section~\ref{evaluation}) the programmer sets flag $A$ in entry point $ep1$ that that does not allow entry point $ep2$ ... (TODO).

>>>>

The aim of this phase (see Figure~\ref{fig:whoop} -- A) is to take an input Linux driver written in C, together with an environmental model, and output an over-approximated version of the program in Boogie IVL, a simple imperative language with well-defined verification-focused semantics used as the input to a number of cutting-edge verifiers (e.g. Boogie and \corral). To achieve this translation, \whoop uses three LLVM\footnote{\url{http://llvm.org}}-based tools: Chauffeur\footnote{\url{https://github.com/mc-imperial/chauffeur}}, Clang\footnote{\url{http://clang.llvm.org}} and SMACK\footnote{\url{https://github.com/smackers/smack}}~\cite{rakamaric2014smack}.

@@

\whoop is based on lockset analysis and thus can report false bugs, because a violation of the locking discipline does not always correspond to a real data race (e.g. when lockfree synchronization is used instead of locking). \whoop also uses over-approximation (e.g. when reading from the shared state) and summarization, which can be sources of false positives. Further, the tool does not currently check for dynamically created locks or for locks that are provided by external libraries, although the later could be easily addressed by providing a mechanism for users to declare custom locks.

Another limitation of \whoop is that it is unable to verify cases where a driver is forced to be accessed by a single process at a time. This \emph{single-open device}~\cite{corbet2005linux} mode is enforced by using an atomic counter to indicate device availability: an entry point executes if the counter has value 1, else blocks. Because \whoop performs pair-wise analysis, it over-approximates this counter, and regard two entry points as running concurrently, even if this is not the case. However,~\cite{corbet2005linux} advises against using single-open device mode, as it hinders user ingenuity (e.g. a user might expect that a device can be accessed concurrently for performance). Thus, because using single-open device mode is considered bad practice, we burden the developer with disregarding any related false bugs.

Finally, another limitation of our approach is the need for precise environmental modeling. Enriching \whoop with domain-specific knowledge to increase precision is an ongoing manual effort, but requires expertise in the Linux kernel. We argue that further increasing the precision of our models is orthogonal to the contributions of this paper. Moreover, even if the symbolic lockset analysis can be imprecise, \whoop can still use any race-freedom guarantees to significantly speedup a precise bug-finders, as seen in Sections~\ref{whoop:bugfinding} and~\ref{evaluation}.
